---
title: "Progress Report 2.3.2 Primary Paper"
author: "Charles Bond"
date: "10/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) 
```


# Current Status: *Partial Success!*

##Sequence of what I did:
I was able to set up labdsv functions for nmds: `dsvdist` and `nmds`. The output of the normal `dist` function and vegan `vegandist` functions generate NA values which cause an error in all the various NMDS functions I've tried to use. I was hoping the `labdsv` functions would fit together. I did find the `bray/curtis` similarity index is available in `dsvdist`, which is the index the paper used, so I am getting warmer. Still, I get an error that appears realted to NAs:
"`Error in dsvdis(Dia.RelAbun.ext, "bray/curtis", weight = rep(1, ncol(x)),  : NA/NaN/Inf in foreign function call (arg 1)`" 
I believe this means that in argument 1, where the dataset is listed, the function stops because there are NA values? But now I'm confused. Okay. So I used the function `complete.cases()` to look for missing values, and am just now realizing that the last two rows are totally blank. Now I've removed those. YAY, and now the NA error is gone, and the `dsvdist` now gives me a Bray-Curtis dissimilarity matrix! Yay! AND THE NMDS FUNCTION RUNS. Okay, see plot below


## The Data:


```{r readdatta}
library(tidyverse)
Diatom_RelAbun_clean<-read.csv("Primary Paper/data1/Diatoms_RelAbun_REDUCED_PRIMER.csv")
Dia.RelAbun.ext<-Diatom_RelAbun_clean %>%
  dplyr::select(-c(Sample))

```

## Packages
```{r}
library("MASS")
library("vegan")
library("labdsv")
x=Dia.RelAbun.ext
dis_dia<- dsvdis(x,"bray/curtis",weight=rep(1,ncol(x)),step=0.0,diag=FALSE, upper=FALSE)
Diatom_NMDS<-nmds(d=dis_dia,k=2,cmdscale(dis_dia,k=2))
plot(Diatom_NMDS)
  
```
  nmds(dis_dia,k=2,)

## Background on the Article  
For 2 years before I came to Tulane I worked in the Periphyton Lab at FIU, where a Ph.D. student/postdoc in our lab was part of a big mesocosm experiment looking at the responses of diatoms, among other things, to saltwater intrusion in the Florida Everglades (Mazzei et. al. 2018). Some of the wetland/soil microbial folks in our class may also find this interesting, the experiment had soil carbon and bacteria community components published in different papers. (I did not participate in this paper or this kind of data analysis).


## References
[1] Mazzei, V., Gaiser, E., Kominoski, E., Wilson, J., Servais, S., Bauman, B., . . . Troxler, F. (2018). Functional and Compositional Responses of Periphyton Mats to Simulated Saltwater Intrusion in the Southern Everglades. Estuaries and Coasts, 41(7), 2105-2119.